---
title: 'Reproducible Research: Peer Assessment 1'
author: "Jianling Peng"
date: "April 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2); library(dplyr)
```

### set up working directory
```{r workingdirectory}
setwd("C:/Users/jpeng11/coursera/Reproducilbe Research/RepData_PeerAssessment1")
```
## Downlodad data
```{r, echo = T, results ='hide'}

unzip(zipfile = "activity.zip", exdir = ".", unzip = "internal")
```


## Loading and preprocessing the data

```{r, echo = T, results = 'hide'}
activity <- read.csv("activity.csv", stringsAsFactors = F, header = T)
head(activity)
activity$Date <- as.Date(activity$date, format = "%Y-%m-%d")
```


## Histogram of the total number of steps taken each day
```{r}
daily_steps <- tapply(activity$steps, activity$Date, sum, na.rm = T)
hist(daily_steps, xlab = "Steps", main = "Histogram of Daily Total Steps")
```

## What is mean and median number of steps  taken per day?
```{r}
mean_step <- mean(daily_steps, na.rm = T)
median_step <- median(daily_steps, na.rm = T)
```

#### The mean steps of total number of steps taken per day is `r mean_step` 
#### The meidan steps of total number of steps taken per day is `r median_step`

## Time series plot of the average number of steps taken
```{r}

intervals <- activity %>% group_by(interval) %>%
    summarise(average = mean(steps, na.rm = T))
gg <- ggplot(data = intervals, aes(x = interval, y = average)) + geom_line() 
gg + labs(y = "Number of Steps")

max_step <- intervals %>% filter(average == max(average))
max_interval <- max_step$interval

```
#### the interval `r max_interval`, on average across all the days in the dataset, contain the maximum number of steps, which is `r max_step$average`.

## Total number of missing values

```{r}
miss_num <- sum(!complete.cases(activity))
```
#### Total number of missing values in the dataset is `r miss_num`.

## Imputing missing values with mean value for each interval

```{r}
imputed_activity <- activity
for (i in 1:dim(imputed_activity)[1]) {
    
    if (is.na(imputed_activity$steps[i])){
        int <- imputed_activity$interval[i]
        sub <-intervals[intervals$interval == int,]$average
        imputed_activity$steps[i] <- sub
    }
    #else imputed_activity$steps <- imputed_activity$steps
}
```

## Total number of steps take per day after imputing missing values

```{r}

daily_steps_imp <- tapply(imputed_activity$steps, imputed_activity$Date, sum, na.rm = T)
hist(daily_steps_imp,xlab = "Steps", main = "Histogram of Daily Total Steps after Imputation")
mean_step <- mean(daily_steps_imp)
median_step <- median(daily_steps_imp)
```

#### The mean steps of total number of steps taken per day afterimputation  is `r mean_step` 
#### The meidan steps of total number of steps taken per day after imputaion is `r median_step`


## Are there differences in activity patterns between weekdays and weekends?

```{r}
imputed_activity <- imputed_activity %>% 
    mutate(wDay = ifelse((weekdays(Date) %in% c("Saturday", "Sunday")),
                         "WEEKEND", "WEEKDAY")) %>% 
    group_by(wDay, interval) %>% 
    summarise(average= mean(steps))

gg <- ggplot(imputed_activity, aes(x = interval, y = average)) + 
    geom_line() + facet_grid(wDay ~.)
gg + labs(y = "Average Number of Steps", title = "Average Number of Steps per  5-min Interval" )

```
